---
title: "National Powerhouses"
subtitle: "Comparing swimmer body profiles across the top ten Olympic swimming nations (1980–2016)"
format:
  html:
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(plotly)

top_teams <- c(
  "United States",
  "Australia",
  "Canada",
  "Great Britain",
  "Japan",
  "Germany",
  "Netherlands",
  "China",
  "Hungary",
  "France"
)

df <- read_csv("data/olympics.csv.gz", show_col_types = FALSE) |>
  filter(
    sport == "Swimming",
    year >= 1980,
    !is.na(height),
    !is.na(weight),
    team %in% top_teams
  ) |>
  mutate(
    bmi = round(weight / (height / 100)^2, 1),
    medal_status = if_else(is.na(medal), "Competitor", paste(medal, "Medalist"))
  )
```

Hover over any swimmer to inspect their name, event, medal result, and physical measurements. Double-click any country in the legend to isolate it, or single-click to compare specific nations.

```{r}
#| label: nation-scatter
#| fig-width: 9
#| fig-height: 6

p <- df |>
  ggplot(aes(
    x = height,
    y = weight,
    color = team,
    text = paste0(
      "<b>", name, "</b> (", team, ")<br>",
      "Event: ", event, "<br>",
      "Year: ", year, " (", city, ")<br>",
      "Status: ", medal_status, "<br>",
      "Height: ", height, " cm | Weight: ", weight, " kg<br>",
      "BMI: ", bmi
    )
  )) +
  geom_point(alpha = 0.65, size = 1.8) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Swimmer Height and Weight by Nation (1980–2016)",
    x = "Height (cm)",
    y = "Weight (kg)",
    color = "Nation"
  )

ggplotly(p, tooltip = "text")
```

## National Comparisons

Traditional powerhouses like the **United States** and **Australia** show remarkably similar distributions, spanning tall sprinters and lean distance athletes across events. Meanwhile, nations like **Japan** have consistently produced medalists with a slightly more compact median stature, highlighting how stroke technique and high cadence can excel at the Olympic level.
